SPANG: A SPARQL Client with Templates, One-Liners, and Programming Interfaces

### 2020-09-29 paper.tex
modify text

\title{SPANG: A SPARQL Client with One-Liners, Template Libraries, and Programming Interfaces}
\titlerunning{SPANG: A SPARQL Client with One-Liners, Template Libraries}
\author{Hirokazu Chiba\orcidID{0000-0003-4062-8903}}
\authorrunning{H. Chiba}
\institute{Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
}
%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
%How can we maximize the value of accumulated SPARQL queries? 
SPARQL is a key component of the Semantic Web, and thus, the reusability of SPARQL is crucial for maximizing productivity on the Semantic Web.
While an increasing number of datasets have been published in the Resource Description Framework (RDF) and the SPARQL standard allows for an interoperable framework for querying the RDF datasets, the reuse of SPARQL queries remains limited. We present SPANG, a framework to make SPARQL queries more reusable using a templating engine, one-liner functionality, and an application programming interface. Here we demonstrate its use cases for making SPARQL queries more reusable; assigning a set of metadata for annotating a SPARQL query, which enables to identify SPARQL query and provides useful information for the query; parameterizing SPARQL queries, which makes each query reusable under different settings; assigning a globally unique and stable identifier for each query; and constructing a server to provide indices of SPARQL queries. SPANG makes SPARQL queries more reusable in the local system or across the Web and leads to the construction of an eco-system wherein SPARQL queries are shared to maximize the productivity of Semantic Web developers and users. SPANG is available at \url{https://spang.dbcls.jp/}.
\blfootnote{Copyright © 2020 for this paper by its author. Use permitted under Creative Commons License Attribution 4.0 International (CC BY 4.0).}
\keywords{SPARQL \and template engine \and one-liner \and query library}

@@

An increasing number of datasets have been published in the Resource Description Framework (RDF).
SPARQL is the standardized interface to RDF and contributing effective reuse of distributed RDF data. 
Although eco-systems for developing SPARQL have been developed, writing SPARQL codes is often a burden for Semantic Web developers and users. 
SPARQL queries have been written in many projects and are the result of developers' efforts. 
How can we maximize the value of accumulated SPARQL queries?
Reuse of the queries is the key to this problem.
For the effective construction of Semantic Web applications, the reuse of such queries is invaluable.
% Reusability is the key to maximize the value of accumulated open data and software.
However, owing to the limitation of SPARQL specifications, such as the lack of a parameterization mechanism, the reuse of these queries remains limited. 
% Reusing SPARQL queries are important for efficient development of the Semantic Web.

We present SPANG and its use cases to make SPARQL queries more reusable, thereby maximizing the value of the written SPARQL queries;
assign metadata to SPARQL queries;
parameterize the SPARQL queries;
assign global identifiers; and
construct a server for providing queries.
%Further, we demonstrate the usefulness of the constructed SPANG template library, which is reusable on the Semantic Web platform.

@@

The SPANG~\cite{spang} project was commenced with the objective of maximizing the usability of SPARQL queries. 
SPANG was originally developed as a command-line SPARQL client~\cite{spang}, and it is now re-implemented in JavaScript using Node.js. Thus, it can not only be called in the command line but also in the JavaScript code used to construct a website.

@@

SPARQL templating mechanisms have been implemented in various projects. Based on the templating mechanisms, SPARQL libraries have been implemented, such as Bioqueries~\cite{bioqueries}. Bioqueries have tried to collect SPARQL queries as community effort. It increased the findability and reusability.
However, the query can only be executed on the web site.
Users should copy and paste the queries into their own software.
In contrast, we attempted to build a software network on top of the Semantic Web platform.

>>>>

\title{SPANG: A SPARQL Client with Templates, One-Liners, and Programming Interfaces}
\titlerunning{SPANG: A SPARQL Client with Templates, One-Liners, and API}
\author{Hirokazu Chiba\orcidID{0000-0003-4062-8903}}
\authorrunning{H. Chiba}
\institute{Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
}
%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
%How can we maximize the value of accumulated SPARQL queries? 
SPARQL is a key component of the Semantic Web, and its reusability of is crucial for maximizing productivity on the Semantic Web.
While an increasing number of datasets have been published in the Resource Description Framework (RDF) and the SPARQL standard allows for an interoperable framework for querying the RDF datasets, the reuse of SPARQL queries remains limited. Herein, we present SPANG, a framework used to facilitate the reuse of SPARQL queries by using a template engine, one-liner functionalities, and programming interfaces. We demonstrate its use for making SPARQL queries more reusable; assign a set of metadata for annotating a SPARQL query, which enables the identification of the SPARQL query and provides useful information for the query; parameterize SPARQL queries, which makes each query reusable under different settings; assign a globally unique and stable identifier for each query; and construct a server for providing the indices of SPARQL queries. SPANG makes SPARQL queries more reusable in the local system or across the Web and facilitate the construction of an eco-system wherein SPARQL queries are shared to maximize the productivity of Semantic Web developers and users. SPANG is available at \url{https://spang.dbcls.jp/}.
\blfootnote{Copyright © 2020 for this paper by its author. Use permitted under Creative Commons License Attribution 4.0 International (CC BY 4.0).}
\keywords{SPARQL \and template engine \and one-liner \and query library}

@@

An increasing number of datasets have been published in the Resource Description Framework (RDF).
SPARQL is the standardized interface to RDF and contributes to the effective reuse of distributed RDF data. 
Although eco-systems have been developed for effective use of SPARQL, writing SPARQL codes is often a burden for Semantic Web developers and users. 
SPARQL queries have been written in many projects and are the result of developers' efforts. 
%How can we maximize the value of accumulated SPARQL queries?
The value of the SPARQL queries accumulated thus far can be maximized by reusing SPARQL queries.
%Reuse of the queries is the key to this problem.
For the effective construction of Semantic Web applications, the reuse of such queries is invaluable.
% Reusability is the key to maximize the value of accumulated open data and software.
However, owing to the limitation of SPARQL specifications, such as the lack of a parameterization mechanism, the reuse of these queries remains limited. 
% Reusing SPARQL queries are important for efficient development of the Semantic Web.

We present SPANG and demonstrate its use for making SPARQL queries more reusable, thereby maximizing the value of the written SPARQL queries;
assign metadata to SPARQL queries;
parameterize the SPARQL queries;
assign global identifiers; and
construct a server for providing queries.
%Further, we demonstrate the usefulness of the constructed SPANG template library, which is reusable on the Semantic Web platform.

@@

The SPANG project was commenced with the objective of maximizing the usability of SPARQL. 
SPANG was originally developed as a command-line SPARQL client~\cite{spang}, and it is now re-implemented in JavaScript using Node.js. Thus, it can not only be called in the command line but also in the JavaScript code used to construct a website.

@@

SPARQL templating mechanisms have been implemented in various projects. SPARQL libraries, such as Bioqueries~\cite{bioqueries}, have been implemented based on the templating mechanisms. Bioqueries comprise an effort to collect SPARQL queries as a community attempt. It contributed the increased findability of queries.
However, the users should still copy and paste the queries into their own software in specific templating formats to reuse them.
In this paper, we present SPANG framework, where we attempt to build a software network with programming interfaces based on the Semantic Web platform.
